---
title: "Netdis: step by step"
author: "Martin O'Reilly, Jack Roberts, Luis Ospina-Forero"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Netdis: Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options: 
chunk_output_type: console

---

```{r, include = FALSE}
  knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  )
```

# Introduction
This Vignette explains some inner calculations of Netdis, which the user doesn't need to go through, however we explain them here for the users that may want to use Netdis in a Modular fashion.

For a simple Netdis function call see instead ["Simple and quick (default) usage 1: pairwise comparisons"](default_pairwise_usage.html).

# What is Netdis?
(Extracted from Ali et al. (2014)): Netdis counts small subgraphs  $w$ on $k$ nodes for all 2-step ego-networks, $k=3,4,5$. These counts are centred by subtracting the expected number of counts $E_w$. These centred counts of each network are then compared thus leading to the Netdis statistic.

Netdis is constructed as follows:

Let $N_{w,i}(G)$ be the number of induced occurrences of small graphs $w$ in the 2-step ego network of vertex $i$. Now, bin all 2-step ego-networks of network $G$ according to their network density. Let $E_w(G,d)$ be the expected number of occurrences of $w$ in an ego-network whose density falls in density bin $d$. For a given network $G$ compute the centred subgraph counts as
\[
S_w(G)=\sum\limits_{i }{\bigg (N_{w,i}(G)- E_w(G, \rho(i)) \bigg )},
\] 
where $i$ is a node in $G$ and $\rho(i)$ the density bin of the 2-step ego-network of node $i$.

Now, to compare networks $G_1$ and $G_2$, set
$$
\displaystyle
netD_2^S(k) = \tfrac{1}{ \sqrt{ M(k)} } \sum\limits_{w \in A(k)}
\bigg ({ \tfrac{S_w(G_1)  S_w(G_2)} {\sqrt{S_w(G_1)^2 + S_w(G_2)^2}}
}\bigg ), \quad k=3,4, 5, 
$$
where $A(k)$ is the set of connected subgraphs of size $k$, and where $M(k)$ is a normalising constant so that $netD_2^S(k)\in[-1,1]$. $M(k)$ is equal to
\[
M(k)  = \sum\limits_{w \in A(k)}
\left( \tfrac{ S_w(G_1)^2 }{\sqrt{S_w(G_1)^2 + S_w(G_2)^2}} \right)
\sum\limits_{w \in A(k)}
\left(\tfrac{ S_w(G_2)^2 } {\sqrt{S_w(G_1)^2 + S_w(G_2)^2}}
\right)
.
\]
The corresponding Netdis statistic is defined as
$$Netdis(k)=netd_2^S(k)=\tfrac{1}{2}(1-netD_2^S(k)) \in [0,1].$$
Small values of Netdis suggest higher `similarity' between the networks. By default Netdis uses subgraphs on $k=4$ nodes.


# Load required libraries
```{r}
# Load libraries
library("netdist")
library("igraph")
library("purrr")
```

# Netdis setp by step

## Networks being compared.
Generation of tree-like networks with 400 nodes and 1600 nodes. 

```{r, netwokrs,fig.align='center'}
# Create networks
set.seed(34)
gTree_1 <- igraph::as.undirected( make_tree(n = 20^2,children = 3) )
gTree_2 <- igraph::as.undirected( make_tree(n = 40^2,children = 3) )

plot(gTree_1,vertex.size=0.8,vertex.label=NA)
plot(gTree_2,vertex.size=0.8,vertex.label=NA)
```



## Set Netdis parameters
```{r}
# Maximum graphlet size to calculate counts and netdis statistic for.
max_graphlet_size <- 4

# Ego network neighbourhood size
neighbourhood_size <- 2

# Minimum size of ego networks to consider
min_ego_nodes <- 3
min_ego_edges <- 1

# Ego network density binning parameters
min_bin_count <- 5
num_bins <- 100
```


## Obtain list of ego networks
```{r}
# Get ego networks for query graphs and reference graph
ego_1 <- make_named_ego_graph(gTree_1, 
                              order = neighbourhood_size, 
                              min_ego_nodes = min_ego_nodes, 
                              min_ego_edges = min_ego_edges)

ego_2 <- make_named_ego_graph(gTree_2, 
                              order = neighbourhood_size, 
                              min_ego_nodes = min_ego_nodes, 
                              min_ego_edges = min_ego_edges)
head(ego_1,n=2)
head(ego_2,n=2)
```

## Count the number of nodes and the graphlets in ego networks
```{r}
# Count graphlets for ego networks in query and reference graphs
graphlet_counts_1 <- ego_to_graphlet_counts(ego_networks = ego_1, max_graphlet_size = max_graphlet_size)
graphlet_counts_2 <- ego_to_graphlet_counts(ego_networks = ego_2, max_graphlet_size = max_graphlet_size)

head(graphlet_counts_1)
head(graphlet_counts_2)
```

## Calculate expected graphlet counts in each bin.

Variants of Netdis that the user can consider: 

* Expectation via a gold-standard network (Original proposal).
* Expectation via the Geometric Poisson / Polya-Aeppli approximation.
* No expectation or Expectation zero.

### Expectation via a gold-standard network
For this case the user must provide the gold-standard network of their choosing. This network will be used as a comparison reference point by Netdis.

```{r,fig.align='center'}
# A gold-standard network
gst_1 <- erdos.renyi.game(n = 30^2,p.or.m = graph.density(graph = gTree_2))
plot(gst_1,vertex.size=0.8,vertex.label=NA)
```

### Obtain ego-network counts and binning according to the edge-density of the gold-standard network ($N_w$ and $\rho(.)$).
```{r}
# Obtain graphlet counts and binning for gold-standard
ego_gst_1 <- make_named_ego_graph(graph = gst_1, 
                              order = neighbourhood_size, 
                              min_ego_nodes = min_ego_nodes, 
                              min_ego_edges = min_ego_edges)

graphlet_counts_gst_1 <- ego_to_graphlet_counts(ego_networks = ego_gst_1,
                                                max_graphlet_size = max_graphlet_size)

head(graphlet_counts_gst_1)

densities_gst_1<- ego_network_density(graphlet_counts = graphlet_counts_gst_1)

# Adaptively bin ego-network densities
binned_densities_gst_1 <- binned_densities_adaptive(densities = densities_gst_1, 
                                                min_counts_per_interval = min_bin_count, 
                                                num_intervals = num_bins)

str(binned_densities_gst_1)
```

### Obain the scaled mean counts for each density bin for the gold-standard network (canonical $E_w$)
```{r}
# Scale ego-network graphlet counts by dividing by total number of k-tuples in
# ego-network (where k is graphlet size)
scaled_graphlet_counts_ref <- scale_graphlet_counts_ego(graphlet_counts = graphlet_counts_gst_1,
                                                        max_graphlet_size =max_graphlet_size)

# Average ref graphlet counts across density bins
ref_binned_graphlet_counts <- mean_density_binned_graphlet_counts(graphlet_counts = scaled_graphlet_counts_ref, density_interval_indexes = binned_densities_gst_1$interval_indexes)

head(ref_binned_graphlet_counts)
```


## Obtain the centred counts for each graphlet on the query graphs based on the GoldStandard counts ($N_w(G) - E_w(G)$).

```{r}
# Scale the reference counts of the gold-standard network to the sizes of each of the query networks.
exp_graphlet_counts_1 <- netdis_expected_counts(graphlet_counts = graphlet_counts_1,
                                                density_breaks = binned_densities_gst_1$breaks, 
                                                density_binned_reference_counts =  ref_binned_graphlet_counts, 
                                                max_graphlet_size = max_graphlet_size,
                                                scale_fn=count_graphlet_tuples)


exp_graphlet_counts_2 <- netdis_expected_counts(graphlet_counts = graphlet_counts_2,
                                                density_breaks = binned_densities_gst_1$breaks, 
                                                density_binned_reference_counts =  ref_binned_graphlet_counts, 
                                                max_graphlet_size = max_graphlet_size,
                                                scale_fn=count_graphlet_tuples)

# Centre graphlet counts by subtracting expected counts
centred_graphlet_counts_1 <- netdis_subtract_exp_counts(graphlet_counts = graphlet_counts_1,
                                                        exp_graphlet_counts = exp_graphlet_counts_1, 
                                                        max_graphlet_size = max_graphlet_size)

centred_graphlet_counts_2 <- netdis_subtract_exp_counts(graphlet_counts = graphlet_counts_2,
                                                        exp_graphlet_counts = exp_graphlet_counts_2, 
                                                        max_graphlet_size = max_graphlet_size)

head(centred_graphlet_counts_1)
head(centred_graphlet_counts_2)
```


## Sum centred graphlet counts across all ego networks ($S_w(G)$)
```{r}
sum_graphlet_counts_1 <- colSums(centred_graphlet_counts_1)
sum_graphlet_counts_1

sum_graphlet_counts_2 <- colSums(centred_graphlet_counts_2)
sum_graphlet_counts_2
```

## Calculate netdis statistics
```{r}

netdis_result <- netdis_uptok(centred_graphlet_count_vector_1 = sum_graphlet_counts_1,
                              centred_graphlet_count_vector_2 = sum_graphlet_counts_2, 
                              max_graphlet_size =  max_graphlet_size)

print(netdis_result)
```
























































