---
title: "Netdis - 2 graphs with Expected Counts from Geometric Poisson Approximation"
author: "Martin O'Reilly, Jack Roberts"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Netdis - 2 graphs with GP Approximation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Load required libraries
```{r}
# Load libraries
library("netdist")
library("purrr")
```

## Load graphs
```{r}
# Set source directory for Virus PPI graph edge files
source_dir <- system.file(file.path("extdata", "VRPINS"), package = "netdist")

# Load query graphs
graph_1 <- read_simple_graph(file.path(source_dir, "EBV.txt"),
                             format = "ncol")

graph_2 <- read_simple_graph(file.path(source_dir, "ECL.txt"),
                             format = "ncol")

```

## Set Netdis parameters
```{r}
# Maximum graphlet size to calculate counts and netdis statistic for.
max_graphlet_size <- 4

# Ego network neighbourhood size
neighbourhood_size <- 2

# Minimum size of ego networks to consider
min_ego_nodes <- 3
min_ego_edges <- 1

# Ego network density binning parameters
min_bin_count <- 5
num_bins <- 100
```

## Generate ego networks
```{r}
# Get ego networks for query graphs and reference graph
ego_1 <- make_named_ego_graph(graph_1, 
                              order = neighbourhood_size, 
                              min_ego_nodes = min_ego_nodes, 
                              min_ego_edges = min_ego_edges)

ego_2 <- make_named_ego_graph(graph_2, 
                              order = neighbourhood_size, 
                              min_ego_nodes = min_ego_nodes, 
                              min_ego_edges = min_ego_edges)

```

## Count graphlets in ego networks
```{r}
# Count graphlets for ego networks in query and reference graphs
graphlet_counts_1 <- ego_to_graphlet_counts(ego_1, max_graphlet_size = max_graphlet_size)
graphlet_counts_2 <- ego_to_graphlet_counts(ego_2, max_graphlet_size = max_graphlet_size)

```

## Bin ego networks by density
```{r}

# Get ego-network densities
densities_1 <- ego_network_density(graphlet_counts_1)
densities_2 <- ego_network_density(graphlet_counts_2)

# Adaptively bin ego-network densities
binned_densities_1 <- binned_densities_adaptive(densities_1, 
                                                min_counts_per_interval = min_bin_count, 
                                                num_intervals = num_bins)

ego_density_bins_1 <- binned_densities_1$breaks

binned_densities_2 <- binned_densities_adaptive(densities_2, 
                                                min_counts_per_interval = min_bin_count, 
                                                num_intervals = num_bins)

ego_density_bins_2 <- binned_densities_2$breaks
```

## Calculate expected graphlet counts in each bin using geometric poisson approximation
```{r}

density_binned_counts_gp <- function(graphlet_counts, bin_indexes, max_graphlet_size) {
  
  mean_binned_graphlet_counts <- mean_density_binned_graphlet_counts(
                                  graphlet_counts, 
                                  bin_indexes)
  
  exp_counts_bin <- function(bin_idx) {
    counts <- graphlet_counts[bin_indexes == bin_idx, ]
    means <- mean_binned_graphlet_counts[bin_idx,]
    
    mean_sub_counts <- sweep(counts, 2, means)
    
    Vd_sq <- colSums(mean_sub_counts^2)/(nrow(mean_sub_counts)-1)
    theta_d <- 2*means / (Vd_sq + means)

    exp_counts_dk <- vector()
    for (k in 2:max_graphlet_size) {
      graphlet_idx <- graphlet_ids_for_size(k)
      
      lambda_dk <- (1 / length(graphlet_idx)) * 
                   sum(
                     2 * means[graphlet_idx]^2 /
                     (Vd_sq[graphlet_idx] + means[graphlet_idx])
                   )
      
      exp_counts_dk <- append(exp_counts_dk,
                              lambda_dk / theta_d[graphlet_idx])
    }
    
    exp_counts_dk
  }
  
  nbins <- length(unique(bin_indexes))
  expected_counts_bin <- t(mapply(exp_counts_bin, bin_idx = 1:nbins))
  
  # deal with NAs caused by bins with zero counts for a graphlet
  expected_counts_bin[is.nan(expected_counts_bin)] = 0
  
  expected_counts_bin
}

binned_graphlet_counts_1 <- density_binned_counts_gp(graphlet_counts_1,
                                                     binned_densities_1$interval_indexes,
                                                     max_graphlet_size)

binned_graphlet_counts_2 <- density_binned_counts_gp(graphlet_counts_2,
                                                     binned_densities_2$interval_indexes,
                                                     max_graphlet_size)

```

## Centre graphlet counts of query graphs using binned expected counts 
```{r}
# Calculate expected graphlet counts for each ego network
exp_graphlet_counts_1 <- netdis_expected_graphlet_counts_per_ego(graphlet_counts_1, 
                                                                 ego_density_bins_1, 
                                                                 binned_graphlet_counts_1,
                                                                 max_graphlet_size,
                                                                 scale_fn = NULL)


exp_graphlet_counts_2 <- netdis_expected_graphlet_counts_per_ego(graphlet_counts_2, 
                                                                 ego_density_bins_2, 
                                                                 binned_graphlet_counts_2,
                                                                 max_graphlet_size,
                                                                 scale_fn = NULL)
# Centre graphlet counts by subtracting expected counts
centred_graphlet_counts_1 <- netdis_subtract_exp_counts(graphlet_counts_1,
                                                            exp_graphlet_counts_1,
                                                            max_graphlet_size)

centred_graphlet_counts_2 <- netdis_subtract_exp_counts(graphlet_counts_2,
                                                        exp_graphlet_counts_2,
                                                        max_graphlet_size)
```

## Sum centred graphlet counts across all ego networks
```{r}
sum_graphlet_counts_1 <- colSums(centred_graphlet_counts_1)

sum_graphlet_counts_2 <- colSums(centred_graphlet_counts_2)
```

## Calculate netdis statistics
```{r}

netdis_result <- netdis_uptok(sum_graphlet_counts_1, 
                              sum_graphlet_counts_2, 
                              max_graphlet_size)

print(netdis_result)
```