---
title: "Quick start guide for Netdis"
author: "Martin O'Reilly"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick start for Netdis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Virus PPI example for Netdis

### Load graphs
Use `read_simple_graphs` to read graph data from all files in a directory that 
match a specific filename pattern in a format suitable for calculating 
graphlet-based feature counts using the 
[ORCA package](https://CRAN.R-project.org/package=orca).
We use `igraph::read_graph` to read graph data from files, so support
all file formats it supports. See help for `igraph::read_graph` for a list of 
supported values for the `format` parameter and the [igraph documentation](http://igraph.org/c/doc/igraph-Foreign.html#igraph_read_graph_edgelist)
for descriptions of each of the supported file formats.

The ORCA package we use to efficiently calculate graphlet and orbit counts 
requires that graphs are _undirected_, _simple_ (i.e. have no self-loops or 
multiple edges) and _connected_ (i.e. have no isolated vertices). Therefore, by
default, graphs loaded by `read_simple_graphs` will be coerced to have the above
properties. This can be avoided by setting the relevant `as_undirected`, 
`remove_loops`, `remove_multiple` or `remove_isolates` parameters to `FALSE`.
```{r}
library("netdist")
edge_format = "ncol"
# Load reference graph (used for Netdis. Not required for NetEMD)
ref_path = file.path(system.file(file.path("extdata", "random"), package = "netdist"), "ER_1250_10_1")
ref_graph <- read_simple_graph(ref_path, format = edge_format)

# Set source directory and file properties for Virus PPI graph edge files
source_dir <- system.file(file.path("extdata", "VRPINS"), package = "netdist")
edge_format = "ncol"
file_pattern = "*"

# Load all graphs in the source folder matching the filename pattern
query_graphs <- read_simple_graphs(source_dir, format = edge_format, 
                             pattern = file_pattern)
print(names(query_graphs))
```

In this example we will use counts of graphlets containing up to 4 nodes and 
consider ego-network neighbourhoods of size 2 (i.e. the immediate neighbours of 
each node plus their immediate neighbours). 
```{r}
# Set the maximum graphlet size to compute counts for
max_graphlet_size <- 4
neighbourhood_size <- 2
```

## Generate a function to generate expected graphlet counts
Use `netdis_expected_graphlet_counts_ego_fn` to generate a function that 
calculates expected ego-network graphlet counts for query graphs based on the 
statistics of a provided reference graph.

```{r}
expected_count_fn <- netdis_expected_graphlet_counts_ego_fn(
  ref_graph, max_graphlet_size, neighbourhood_size) 
```

## Generate centred graphlet counts for a set of query graphs
```{r}
centred_counts <- purrr::map(query_graphs, netdis_centred_graphlet_counts,
                             max_graphlet_size = max_graphlet_size, 
                             neighbourhood_size = neighbourhood_size,
                             expected_ego_count_fn = expected_count_fn)
```

## Generate NetDis measures between each pair of query graphs
```{r}
# Netdis measure for graphlets of size 3
res3 <- netdis_for_all_graphs(centred_counts, 3)
netdis3_mat <- cross_comp_to_matrix(res3$netdis, res3$comp_spec)
# Netdis measure for graphlets of size 4
res4 <- netdis_for_all_graphs(centred_counts, 3)
netdis4_mat <- cross_comp_to_matrix(res4$netdis, res4$comp_spec)
netdis4_mat
```

## Generate dendrograms
```{r}
graphdists<-as.dist(netdis4_mat)
par(mfrow=c(1,2))
cex=1
# Dendrogram based on Netdis measure for graphlets of size 3
title = paste("Netdis: graphlet size = ", 3, sep = "")
plot(phangorn::upgma(as.dist(netdis3_mat), method="average"), use.edge.length=FALSE, 
     edge.width=cex*2, main=title, cex.lab=cex, cex.axis=cex, cex.main=cex, 
     cex.sub=cex, cex=cex)
# Dendrogram based on Netdis measure for graphlets of size 4
title = paste("Netdis: graphlet size = ", 4, sep = "")
plot(phangorn::upgma(as.dist(netdis4_mat), method="average"), use.edge.length=FALSE, 
     edge.width=cex*2, main=title, cex.lab=cex, cex.axis=cex, cex.main=cex, 
     cex.sub=cex, cex=cex)
```
```{r}
cex=1.5
col <- colorRampPalette(colors = c("blue","white"))(100)
title = paste("Netdis: graphlet size = ", 3, sep = "")
heatmap(netdis3_mat, Rowv = NULL, Colv = NULL, col = col, main = title, cexRow = cex, cexCol = cex, symm = TRUE)
```
```{r}
cex=1.5
col <- colorRampPalette(colors = c("blue","white"))(100)
title = paste("Netdis: graphlet size = ", 4, sep = "")
heatmap(netdis4_mat, Rowv = NULL, Colv = NULL, col = col, main = title, cexRow = cex, cexCol = cex, symm = TRUE)
```